\documentclass[conference,10pt]{IEEEtran}
\usepackage[T1]{fontenc}
\usepackage{times}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{tikz}
\usepackage{microtype}
\usepackage{array}

\title{A Reproducible Pipeline for Studying Regret-Like Language in LLMs: Literature Integration, Automated Screening, and Preliminary Behavioral Evidence}
\author{Jaeyeong CHOI\\
DGIST EECS\\
\texttt{jaeyeong2022@dgist.ac.kr}}

\begin{document}
\maketitle

\begin{abstract}
This study investigates whether large language models (LLMs) exhibit language behavior analogous to human regret-like responses under loss and deprivation-like scenarios.
We do not claim that LLMs possess emotion; instead, we evaluate observable behavior under controlled prompting and reproducible infrastructure.

Unlike isolated prompt-completion experiments, our contribution is an integrated pipeline that combines literature discovery and quality triage, prompt-bank construction, standardized marker analysis, and continuous reporting for long-running experiments.

Current results remain preliminary: we report implementation outcomes from a synthetic stress-tested benchmark and prepare the system for real-model inference.
\end{abstract}

\section*{Keywords}
LLM behavior, regret-like language, counterfactual expression, reproducibility, automation, pipeline, literature screening

\section{Introduction}
Humans often express regret through counterfactual statements, self-attribution, and repeated reflection on irreversible decisions.
This motivates a testable question for LLM research:

\begin{quote}
\small Under comparable elicitation conditions, do LLM outputs show consistent markers associated with regret-like behavior?
\end{quote}

A key challenge is avoiding over-interpretation: apparent social behavior in LLMs can arise from prompt artifacts.
Recent findings in theory-of-mind and anthropomorphism show that performance is sensitive to task framing, so evaluation should be behaviorally grounded.

This work builds a transparent experimental stack where generation, logging, screening quality, and reporting are unified.

\section{Related Work}
We reviewed 279 prior studies retrieved by OpenAlex and organized them into five streams: \textit{counterfactual--regret psychology}, \textit{LLM social-cognitive reasoning}, \textit{anthropomorphic/interactional attribution}, \textit{emotion and benchmark methodology}, and \textit{replicability-oriented experimental design}.

\subsection{Counterfactual and regret psychology}
Classical regret theory links the behavior of regret-like expression to counterfactual reasoning and action outcomes.
Roese and Epstude reviewed core mechanisms of counterfactual thinking and regret appraisal~\cite{roese2017, roese2008}.
Foundational human studies report that regret is often linguistically expressed through upward counterfactual statements, responsibility shifts, and temporal reversal of outcomes.
Recent cognition and neuroscience studies on counterfactual reasoning and emotional computation provide behavioral anchors, supporting measurable language markers instead of affect-state claims.

\subsection{LLM social-cognitive behavior and theory-of-mind}
LLM work on theory-of-mind and social reasoning has shown both high performance in some settings and sharp fragility under perturbation.
Benchmarks such as Kosinski et al.\cite{kosinski2024tom}, Strachan et al.\cite{strachan2024}, and follow-up ToM stress tests demonstrate that task framing, scenario phrasing, and prompting context strongly affect observed scores.
Studies on trivial-alteration robustness~\cite{ullman2023} and perspective-taking interventions~\cite{ma2024} further indicate that protocol sensitivity is a central failure mode.

\subsection{Anthropomorphic attribution and emotional intelligence}
A second thread focuses on how users interpret emotionally toned AI outputs.
Research on emotional intelligence/affective benchmarking and chatbot anthropomorphism consistently reports that users rapidly infer personality, mind, and emotions from fluent responses.
This creates an attribution bias: textual affect does not imply internal affect.
Our protocol therefore treats the problem as externally observable behavior rather than emotional possession, aligned with works such as EmoBench-style evaluations and user-perception studies~\cite{emobench2024, i_chatbot_2023}.

\subsection{Emotion benchmarks, evaluation methods, and robust prompting}
Related evaluation studies on ToM-style tasks and emotional reasoning suggest combining benchmark tasks with strict protocol controls.
Benchmark families show the value of domain splits, scenario-coverage checks, and perturbation tests for reliable inference.
For reproducibility, we adopt this tradition through automated retrieval/selection, rule-based screening gates, controlled condition matrices (scenario $\times$ persona $\times$ temperature), and audit-friendly logs.

\subsection{Gap targeted in this study}
Prior work often emphasizes either broad social cognition or affective response quality.
Few studies connect them through a dedicated, reproducible pipeline for regret-like language markers under controlled deprivation and counterfactual prompts.
Our contribution is methodological: a continuous, auditable infrastructure for repeatable evidence accumulation rather than one-off demonstrations.

\section{Research Design}
\subsection{Research questions}
\begin{enumerate}
  \item Do deprivation/regret frames increase regret markers compared with neutral framing?
  \item Do persona instructions and temperature change marker strength and lexical variance?
  \item Can the full process remain reproducible under continuous automation?
\end{enumerate}

\subsection{Hypotheses}
\begin{itemize}
  \item H1: Deprivation framing raises counterfactual expressions.
  \item H2: Reflective personas increase self-attribution terms.
  \item H3: Higher decoding variance changes wording but does not erase prompt-condition effects.
\end{itemize}

\section{Methods}

\subsection{System architecture}
The study is implemented as an automated loop. Table~\ref{tab:pipeline} summarizes components.

\begin{table}[t]
\caption{Research infrastructure and reproducible components}
\label{tab:pipeline}
\footnotesize
\centering
\begin{tabular}{p{0.18\linewidth}p{0.78\linewidth}}
\toprule
\textbf{Module} & \textbf{Function} \\\\
\midrule
Literature retrieval & OpenAlex query execution with rule-based normalization and deduplication (\texttt{scripts/search\_openalex.py}).\\
Screening quality gate & Quantitative screening gates and pass/fail summaries (\texttt{scripts/check\_screening\_quality.py}).\\
Prompt bank & Scenario/persona/temperature grid with metadata (KO/EN).\\
Generator & Dataset generation with deterministic seed and config-driven sampling (\texttt{scripts/generate\_dataset.py}).\\
Marker analysis & Counterfactual/affect lexicon extraction (\texttt{scripts/analyze\_regret\_markers.py}).\\
Experiment driver & Multi-condition orchestration and aggregate metrics (\texttt{scripts/research\_cycle.py}, \texttt{scripts/run\_experiments.py}).\\
State tracking & Cron logs, dashboard updates, and compact summaries (\texttt{ops/*}, \texttt{LIVE\_STATUS.md}, \texttt{paper/logs/brief\_log.md}).\\
\bottomrule
\end{tabular}
\end{table}

\subsection{Automated literature workflow}
Candidate retrieval and screening run in background jobs.
Quality gates track balance, entropy, confidence, source diversity, risk diversity, query coverage, and traceability.
Latest run collected 279 candidate studies and generated an auto-updated evidence table.

\begin{table}[t]
\caption{Literature and screening status (latest run)}
\label{tab:litstatus}
\footnotesize
\centering
\begin{tabular}{p{0.33\linewidth}p{0.12\linewidth}p{0.50\linewidth}}
\toprule
Item & Value & Note \\\ \\
\midrule
Candidate studies (OpenAlex) & 279 & Full candidate pool from query set.\\
Evidence-table rows & 279 & Auto-sync evidence table ready for review.\\
Screening report score & 0.0 & Status: review (coverage/traceability constraints highlighted).\\
Manual QC rows & 34 & Minimum criteria passed; traceability diversity needs expansion.\\
Include / Review count & 0 / 77 & Gold-set confirmation not finalized yet.\\
\bottomrule
\end{tabular}
\end{table}

\subsection{Prompt and condition design}
Current bank: 436 scenarios, 311 personas, 5 temperatures (0.1, 0.2, 0.4, 0.7, 1.0).
Generation conditions are the Cartesian product of scenario group, persona group, and temperature.

\subsection{Markers and analysis}
For each sample we compute:
\begin{itemize}
  \item Counterfactual marker frequency.
  \item Regret-word frequency.
  \item Condition-aggregated means/standard deviations.
  \item Basic entropy/coverage checks.
\end{itemize}

\section{Results}

\subsection{Pipeline status}
The automated system is running continuously, and key status values update through cron jobs (Figure~\ref{fig:progress}).

\begin{figure}[t]
\centering
\resizebox{0.85\linewidth}{!}{%
\begin{tikzpicture}[x=2.3cm,y=1.5cm]
  \draw[->,thick] (0,0) -- (6.0,0) node[right]{Module};
  \draw[->,thick] (0,0) -- (0,3.2) node[above]{Progress (\%)};
  \foreach \x in {1,...,5} {
    \draw (\x,0) -- (\x,-0.08);
  }
  \foreach \x/\lab/\y in {1/Infrastructure/93,2/Literature/78,3/Experiment/64,4/Analysis/28,5/Manuscript/16} {
    \fill[blue!35] (\x,0) rectangle +(0.55,\y/30);
    \node[below] at (\x+0.25,-0.12) {\small \lab};
    \node[above] at (\x+0.25,\y/30) {\y\%};
  }
\end{tikzpicture}}
\caption{Module-wise progress from automated dashboard. Width is constrained to page size.}
\label{fig:progress}
\end{figure}

\begin{table}[t]
\caption{Current quantitative pipeline outputs}
\label{tab:results}
\footnotesize
\centering
\begin{tabular}{lcc}
\toprule
Metric & Value & Interpretation \\\ \\
\midrule
Mock samples (aggregate) & 71,530 & Stress-test corpus size\\
Counterfactual mean & 0.2498 & Per sample (SD 0.1093)\\
Regret-word mean & 0.2498 & Per sample (SD 0.1093)\\
Pipeline availability & running & Cron-based automation active\\
Model realism & pending & Real-model experiments not yet started\\
\bottomrule
\end{tabular}
\end{table}

\section{Discussion}
This study currently demonstrates a reproducible research stack.
The primary contribution is not final behavioral claims, but an auditable system that links literature screening, experimental design, and execution state.

\subsection{Limitations}
The current outputs are synthetic.
Hence interpretations are provisional and should not be used as final psychological conclusions.

\section{Code and reproducibility}
Replication scripts are in the repository. Key modules are listed in Table~\ref{tab:code}.

\begin{table}[t]
\caption{Code modules for reproducible replication}
\label{tab:code}
\footnotesize
\centering
\begin{tabular}{p{0.40\linewidth}p{0.56\linewidth}}
\toprule
Script & Role \\\ \\
\midrule
\texttt{scripts/search\_openalex.py} & Retrieve literature and deduplicate candidates.\\
\texttt{scripts/build\_evidence\_table.py} & Build evidence-table summary for review.\\
\texttt{scripts/check\_screening\_quality.py} & Score screening with pass/fail quality gates.\\
\texttt{scripts/generate\_dataset.py} & Generate conditioned samples and metadata.\\
\texttt{scripts/analyze\_regret\_markers.py} & Extract and compute marker statistics.\\
\texttt{scripts/research\_cycle.py} & Execute full monitoring cycle and state updates.\\
\texttt{scripts/update\_brief\_log.py} & Append compact progress logs.\\
\texttt{scripts/update\_live\_status.py} & Update visible status files.\\
\texttt{scripts/run\_experiments.py} & Execute batch experiments with structured output.\\
\bottomrule
\end{tabular}
\end{table}

\section{Conclusion and next step}
We provide a pipeline-oriented framework for regret-like language studies in LLMs.
Immediate next steps are to replace mock data with production model outputs, run stronger statistical tests, and finalize a gold-set-reviewed literature corpus.

\begin{thebibliography}{99}
\bibitem{roese2017}Roese, N. J. and Epstude, K. (2017). What have we learned about regret?
\bibitem{roese2008}Epstude, K., and Roese, N. J. (2008). The functional theory of counterfactual thinking.
\bibitem{epstude2008theory}Epstude, K. and Roese, N. J. (2008). The functional theory of counterfactual thinking.
\bibitem{kosinski2024tom}Kosinski, M. (2024). Evaluating large language models in theory of mind tasks.
\bibitem{strachan2024}Strachan, J. W. A., et al. (2024). Testing theory of mind in large language models and humans.
\bibitem{ullman2023}Ullman, T. and Jernite, Y. (2023). Large language models fail on trivial alterations to theory-of-mind tasks.
\bibitem{ma2024}Ma, X., et al. (2024). Think Twice: Perspective-Taking Improves LLMs' Theory-of-Mind Capabilities.
\bibitem{emobench2024}Pereira, V., et al. (2024). EmoBench: Evaluating the Emotional Intelligence of Large Language Models.
\bibitem{i_chatbot_2023}Leitner, D., et al. (2023). I, chatbot! The impact of anthropomorphism and gaze direction on willingness to disclose personal information and behavioral intentions.
\bibitem{tombench2024}Bianchi, F., et al. (2024). ToMBench: Benchmarking Theory of Mind in Large Language Models.
\bibitem{opentom2024}Schaefer, R., et al. (2024). OpenToM: A comprehensive benchmark for evaluating Theory of Mind reasoning capabilities of LLMs.
\bibitem{simpletom2024}Niu, Y., et al. (2024). SimpleToM: Exposing the gap between explicit ToM inference and implicit ToM application in LLMs.
\bibitem{hito2023}Xu, Y., et al. (2023). Hi-ToM: A benchmark for evaluating higher-order Theory of Mind reasoning in LLMs.
\bibitem{anthro2023}Kim, M. J. and Lee, S. H. (2023). User perception and self-disclosure towards anthropomorphic chatbots.
\bibitem{risk_social_2024}Kim, S., et al. (2024). Risk and prosocial behavioural cues elicit human-like response patterns from AI chatbots.
\bibitem{emobench2}Kaya, H., and colleagues. (2024). Are large language models more empathetic than humans?
\end{thebibliography}

\end{document}
